Ribosome accumulation during early phase resistance training

### 2020-09-03 manuscript.Rmd
updates to fig 1 and 2

load.stats <- read_excel("./data/tr010_training.xlsx", sheet = 1, na = "NA") %>%
  inner_join(read_excel("./data/leg_randomization.xlsx")) %>%
  filter(exercise == "legext") %>%
  mutate(set.load = repetitions * load, 
         week = if_else(session %in% c(1:4), "W1", if_else(session %in% c(5:8), "W2", "W3"))) %>%
  group_by(participant, cond, session, week) %>%
  summarise(load = mean(load, na.rm = TRUE)) %>%
  mutate(week = factor(week, levels = c("W1", "W2", "W3"))) %>%
  print()

@@

>>>>

load.stats <- read_excel("./data/tr010_training.xlsx", sheet = 1, na = "NA") %>%
  inner_join(read_excel("./data/leg_randomization.xlsx")) %>%
  filter(exercise == "legext") %>%
  mutate(set.load = repetitions * load, 
         week = if_else(session %in% c(1:4), "W1", if_else(session %in% c(5:8), "W2", "W3"))) %>%
  group_by(participant, cond, session, week) %>%
  summarise(load = mean(load, na.rm = TRUE)) %>%
  mutate(week = factor(week, levels = c("W1", "W2", "W3")))

@@

```{r total-rna}

# Segmented model analysis in text 
# Read the segmented model 
tc.m2 <- readRDS("./data/derivedData/total-rna-analysis/tc.m2.RDS")

time1.hyp <- hypothesis(tc.m2, "time1 = 0")
time2.hyp <- hypothesis(tc.m2, c("time2 + time1 = 0")) # The slope (on log scale) between session 4 and 8
time3.hyp <- hypothesis(tc.m2, c("time2 + time1 +time3 = 0")) # The slope on log scale between session 8 and 12

slope0_4 <- paste0(sprintf("%.1f", 100 * (exp(time1.hyp$hypothesis[1,2]) - 1)),
                           "%, 95% CI: [",
                   sprintf("%.1f", 100 * (exp(time1.hyp$hypothesis[1,4]) - 1)), 
                   ", ",
                   sprintf("%.1f", 100 * (exp(time1.hyp$hypothesis[1,5]) - 1)),
                   "]")

slope4_8 <- paste0(sprintf("%.1f", 100 * (exp(time2.hyp$hypothesis[1,2]) - 1)),
                           "%, [",
                   sprintf("%.1f", 100 * (exp(time2.hyp$hypothesis[1,4]) - 1)), 
                   ", ",
                   sprintf("%.1f", 100 * (exp(time2.hyp$hypothesis[1,5]) - 1)),
                   "]")

slope8_12 <- paste0(sprintf("%.1f", 100 * (exp(time3.hyp$hypothesis[1,2]) - 1)),
                           "%, [",
                   sprintf("%.1f", 100 * (exp(time3.hyp$hypothesis[1,4]) - 1)), 
                   ", ",
                   sprintf("%.1f", 100 * (exp(time3.hyp$hypothesis[1,5]) - 1)),
                   "]")

Compared to control, a single session of RT led to robust increase in 45S and 47S pre-rRNA (Figure 2B) expressed per unit tissue weight. These were also increased compared to control after twelve sessions together with the 45S ITS and mature rRNA species 28S and 18S. After eight days of rest, only 18S remained elevated above control group  (Figure 2B). This general pattern of rRNA expression was reflected also in total RNA expressed per unit tissue weight were a robust increase compared to contral was seen after session twelve (Figure 2B). Within the experimental group, no differences were seen between volume conditions. Most of the increase was seen in after the first four sessions where the average increase was `r slope0_4` per session in the intervention group as a whole. Between session four and eight, and eight to twelve changes in total RNA per tissue weight per session was not different from zero (`r slope4_8` and `r slope8_12`).
